ODGI: understanding pangenome graphs

### 2021-10-14 manuscript/main.tex
sharing the magic

We store the graph in a vector of node structures, each of which presents a node-local view of the graph sequence, topology, and path layout.
Expressed in terms of the variation graph $V$, the structure for node  the node sequence $seq(n_i)$

>>>>

We store the graph in a vector of node structures, each of which presents a node-local view of the graph sequence, topology, and path layout.
Expressed in terms of the variation graph $V$, ODGI's core $Node$ structure includes a decoder that maps the neighbors of each node to a dense range of integers.
For a given node $i$ and neighbor $j$, the decoder itself does not store the $id$s of $j$, but rather $Node_i.id - Node_j.id$.
This keeps even the size of the encoding small, per common variation graph property (2).

\begin{algorithm}
\MyStruct{Node}{
    \textbf{id} $\in \mathbb{N}$ \tcp{an identifier}
    \textbf{lock} \tcp{atomic locking primitive}
    \textbf{sequence} $= [$A$|$T$|$G$|$C$]*$ \tcp{DNA, RNA, protein}
    \tcp{bit-packed vector of edges}
    \textbf{edges} $= (s_{x},s_{y})* : x \land y \in [1\ldots N]$ \\
    \tcp{bit-packed vector of id deltas}
    \textbf{decoding} $x_i \ldots x_N \in \mathbb{N}^N$ \\
    \tcp{bit-packed vector of path steps}
    \textbf{paths} $[Step_1 \ldots Step_M]*$
}
\MyStruct{Step}{
    \textbf{path\_id} $\in \mathbb{N}$ \tcp{the path's global id}
    \textbf{is\_rev} $\in ( 0, 1 )$ \tcp{the step orientation}
    \textbf{is\_start} $\in ( 0, 1 )$ \tcp{if first step in path}
    \textbf{is\_end} $\in ( 0, 1 )$ \tcp{if last step in path}
    \textbf{prev$_\delta$} $\in [1\ldots N]$ \tcp{$\delta$-encoded previous node}
    \textbf{prev\_rank} $\in \mathbb{N}$ \tcp{step rank on previous node}
    \textbf{next$_\delta$} $\in [1\ldots N]$ \tcp{$\delta$-encoded previous node}
    \textbf{next\_rank} $\in \mathbb{N}$ \tcp{step rank on next node}
}
\caption{ODGI's relativistically-packed node structure and the step structure used to represent the paths as doubly-linked lists.}
\end{algorithm}

%\begin{enumerate}
%\item a node sequence $seq(n_i)$
%\end{enumerate}
